CrypTool Script \\ Mathematics and Cryptography

### 2003-04-01 script/de/chapters/numbertheory.tex
more fixes

\begin{satz}\label{thm-zth-ordp}\footnote{%
    F"ur Primmoduli $p$ haben alle $0 < a < p$ die Ordnung $J(p) = p - 1$.
    Vergleiche dazu Tabelle~5. In diesem Fall nimmt $a^i (\mbox{mod }n)$
    die Werte $1,\dots,p-1$ an. Dieses Aussch"opfen des
    Wertebereiches\index{Wertebereich} ist eine wichtige kryptographische
    Eigenschaft (vergleiche Satz \ref{thm-zth-exhperm}). Hiermit wird eine
    Permutation\index{Permutation} $\pi(p-1)$ festgelegt.
}${}^,$\footnote{%
Tabelle~6 zeigt, dass bei zusammengesetzten Moduli $n$ nicht alle $a$ die
maximale Ordnung $J(n)$ haben. In diesem Beispiel haben nur $5,7,11,15,17,21$
die Ordnung 22.
} F"ur einen Modul $n$ und $a$ teilerfremd zu $n$ gilt: $\{ a^i (\mbox{mod }n) | i =
1,\dots,J(n)\}$ ist gleich der multiplikativen Gruppe $Z_n^*$ genau dann, wenn  $ord_n(a) = J(n)$.
\end{satz}

>>>>

\begin{satz}\label{thm-zth-ordp}\footnote{%
    F"ur Primmoduli $p$ haben alle $a$ mit $0 < a < p$ die Ordnung $J(p) = p
    - 1$.  Vergleiche dazu Tabelle~5. In diesem Fall nimmt $a^i (\mbox{mod
    }n)$ alle Werte $1,\dots,p-1$ an. Dieses Aussch"opfen des
    Wertebereiches\index{Wertebereich} ist eine wichtige kryptographische
    Eigenschaft (vergleiche Satz \ref{thm-zth-exhperm}). Hiermit wird eine
    Permutation\index{Permutation} $\pi(p-1)$ festgelegt.
}${}^,$\footnote{%
Tabelle~6 zeigt, dass bei zusammengesetzten Moduli $n$ nicht alle $a$ die
maximale Ordnung $J(n)$ haben. In diesem Beispiel haben nur $5,7,11,15,17,21$
die Ordnung 22.
} F"ur einen Modul $n$ und $a$ teilerfremd zu $n$ gilt: $\{ a^i (\mbox{mod }n) | i =
1,\dots,J(n)\}$ ist gleich der multiplikativen Gruppe $Z_n^*$ genau dann, wenn  $ord_n(a) = J(n)$.
\end{satz}

### 2003-04-01 script/en/chapters/numbertheory.tex
more fixes

Cryptographically not only the uniqueness of the inverse is important, but also that the
set of possible values is exhausted.
\begin{theorem}\label{thm-zth-exhperm}
For $a,i\in \{1, \cdots, m-1 \}$ with ${\rm~gcd} (a,m)=1),$ then takes the product $a*i {\rm ~mod ~} m $ for a 
certain number $a$ all values from $\{1, \cdots ,m-1 \} $ (exhaustive permutation \index{Permutation} of the Length $m-1$)\footnote{See
also theorem \ref{thm-zth-ordp} in \hyperlink{Kap_3_9}{chapter 3.9 Multiplicative order and primitive roots}.}.
\end{theorem}
\vskip +10 pt
The following three examples\footnote{%
See \hyperlink{AppArith1}{Appendix D} for the source code to compute
the tables using Mathematica\index{Mathematica} and Pari-GP.\index{Pari-GP}
}
illustrate the properties of multiplicative inverses.

@@

\begin{theorem}\label{thm-zth-ordp}\footnote{%
    For prime moduli $p$ all $0 < a < p$ are of order $J(p) = p - 1$.
    Compare table 5 for an example. In this case $a^i (\mbox{mod }n)$ goes
    through the values $1,\dots,p-1$. Exhausting all possible values of the
    set is an important cryptographic proposition (compare theorem
    \ref{thm-zth-exhperm}). This determines a permutation\index{Permutation}
    $\pi(p-1)$.}$^,$\footnote{%
    Table 6 demonstrates that for composite moduli $n$ not all $a$ are of
    maximal order $J(n)$. In this example only $5,7,11,15,17,21$ are of
    order 22.
  } For a modulus $n$ and $a$ relative prime to
  $n$ the following holds: \\
  $\{ a^i (\mbox{mod }n) | i = 1,\dots,J(n)\}$ equals the multiplicative
  group $Z_n^*$ if and only if $ord_n(a) = J(n)$.
\end{theorem}

>>>>

Cryptographically not only the uniqueness of the inverse is important, but also that the
set of possible values is exhausted.
\begin{theorem}\label{thm-zth-exhperm}
For $a,i\in \{1, \cdots, m-1 \}$ with ${\rm~gcd} (a,m)=1),$ then takes the product $a*i {\rm ~mod ~} m $ for a 
certain number $a$ all values from $\{1, \cdots ,m-1 \} $ (exhaustive permutation \index{Permutation} of the length $m-1$)\footnote{See
also theorem \ref{thm-zth-ordp} in \hyperlink{Kap_3_9}{chapter 3.9 Multiplicative order and primitive roots}.}.
\end{theorem}
\vskip +10 pt
The following three examples\footnote{%
See \hyperlink{AppArith1}{Appendix D} for the source code to compute
the tables using Mathematica\index{Mathematica} and Pari-GP.\index{Pari-GP}
}
illustrate the properties of multiplicative inverses.

@@

\begin{theorem}\label{thm-zth-ordp}\footnote{%
    For prime moduli $p$ all $a$ with $0 < a < p$ are of order $J(p) = p - 1$.
    Compare table 5 for an example. In this case $a^i (\mbox{mod }n)$ goes
    through all the values $1,\dots,p-1$. Exhausting all possible values of
    the set is an important cryptographic proposition (compare theorem
    \ref{thm-zth-exhperm}). This determines a permutation\index{Permutation}
    $\pi(p-1)$.}$^,$\footnote{%
    Table 6 demonstrates that for composite moduli $n$ not all $a$ are of
    maximal order $J(n)$. In this example only $5,7,11,15,17,21$ are of
    order 22.
  } For a modulus $n$ and $a$ relative prime to
  $n$ the following holds: \\
  $\{ a^i (\mbox{mod }n) | i = 1,\dots,J(n)\}$ equals the multiplicative
  group $Z_n^*$ if and only if $ord_n(a) = J(n)$.
\end{theorem}
